---
title: ''
author: ""
date: ""
output:
  pdf_document:
    toc: true
    toc_depth: 1
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

\setcounter{page}{2}

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(MASS)
library(knitr)
library(kableExtra)
library(scales)
library(tidyr)
library(RColorBrewer)
```

# Problem Description and Research Question

\textbf{Target audience}: Government and policy makers working on youth
employment \newline

Why do some young people find a job quickly after completing their
education, while others take months or even years to enter the labour
market? \newline

A faster transition into employment is not only beneficial for
individuals, but also for society as a whole. Early labour market
integration means:

\begin{itemize}
\item quicker economic contribution
\item higher productivity and value creation
\item reduced risk of long-term unemployment
\end{itemize}

However, these transitions are not equally distributed. Differences may
arise from individual characteristics, access to work experience, mental
health, or broader structural conditions such as national labour market
systems.

\textbf{Research question:}

> What factors determine how quickly young people enter the labour
> market after completing their education?

To answer this question, an ordinal regression model was estimated using
data from the CRONOS-3 survey, focusing on the duration of active job
search until the first employment.

# Model

Outcome: Time to first job after active search (\textit{w3sq69b})
\newline Input:

\begin{itemize}
\item Birth year (\textit{yrbrn})
\item Country (\textit{cntry})
\item Gender (\textit{gender})
\item Strong Pillar:
\begin{itemize}
\item Years of full-time education completed (\textit{eduyrs})
\item Respondent born in country $\rightarrow$ indicator for migration background (\textit{w2sq18a})
\item Importance of work in life (\textit{w4sq1}: 1$=$Essential, 2$=$Very important, 3$=$Fairly important, 4$=$Not very important, 5$=$Not important at all)
\item Done during studies (\textit{w1sq6\_1} - \textit{w1sq6\_4}: 1$=$Yes, 0$=$No):
\begin{itemize}
\item Studied abroad
\item Internship or work placement abroad
\item Attended language course abroad
\item Summer school, workshop or similar activity abroad
\end{itemize}
\item Apprentice or Internship (\textit{w1sq11\_1}-\textit{w1sq11\_4}: 1$=$Yes, 0$=$No):
\begin{itemize}
\item Participated in unpaid apprenticeship
\item Participated in paid apprenticeship
\item Participated in unpaid internship
\item Participated in paid internship
\end{itemize}
\end{itemize}
\item Health Pillar:
\begin{itemize}
\item Subjective mental health in general (\textit{w3hq57}: 1$=$Very good, 2$=$Good, 3$=$Fair, 4$=$Bad, 5$=$Very bad)
\end{itemize}
\item Equal Pillar:
\begin{itemize}
\item Expect to receive substantial inheritance in future $\rightarrow$ indicator for socioeconomic background (\textit{w4eq10})
\end{itemize}
\end{itemize}

For the fitting of the model, all uninformative categories and missing
values were omitted.

```{r dataprep, include=FALSE, message=FALSE}
# Data ----
## load pillars ----
df_strong = read.csv(file="CRONOS-3 Theme - Make it Strong/CRONOS-3 Theme - Make it Strong.csv")
df_strong$idunique = paste0(df_strong$idno, "_", df_strong$yrbrn, "_", df_strong$cntry)
df_strong = df_strong[, c("idunique", setdiff(names(df_strong), "idunique"))]
non_unique_ids = names(which(table(df_strong$idunique)>1))
nrow(df_strong[df_strong$idunique %in% non_unique_ids,]) # --> not the same people
# View(df_strong)

df_equal = read.csv(file="CRONOS-3 Theme - Make it Equal/CRONOS-3 Theme - Make it Equal.csv")
df_equal$idunique = paste0(df_equal$idno, "_", df_equal$yrbrn, "_", df_equal$cntry)
df_equal = df_equal[, c("idunique", setdiff(names(df_equal), "idunique"))]
non_unique_ids = names(which(table(df_equal$idunique)>1))
nrow(df_equal[df_equal$idunique %in% non_unique_ids,]) # --> not the same people
# View(df_equal)

df_healthy = read.csv(file="CRONOS-3 Theme - Make it Healthy/CRONOS-3 Theme - Make it Healthy.csv")
df_healthy$idunique = paste0(df_healthy$idno, "_", df_healthy$yrbrn, "_", df_healthy$cntry)
df_healthy = df_healthy[, c("idunique", setdiff(names(df_healthy), "idunique"))]
non_unique_ids = names(which(table(df_healthy$idunique)>1))
nrow(df_healthy[df_healthy$idunique %in% non_unique_ids,]) # --> not the same people
# View(df_healthy)

## merge data ----
length(intersect(df_strong$idunique, df_equal$idunique))
length(intersect(df_strong$idunique, df_healthy$idunique))
length(intersect(df_equal$idunique, df_healthy$idunique))

## all the same participants
common_cols = colnames(df_strong)[1:24]

# merge datasets
df = df_strong %>%
  left_join(df_equal[, c("idunique", "w4eq10")], by = "idunique") %>%
  left_join(df_healthy[, c("idunique", "w1hq8", "w3hq57")], by = "idunique")

# rename variables
df = df %>%
  rename(
    timetojob = w3sq69b,
    during_studies_abroad_studies = w1sq6_1,
    during_studies_abroad_work = w1sq6_2,
    during_studies_abroad_langcourse = w1sq6_3,
    during_studies_abroad_summerschool = w1sq6_4,
    workimportance_subj = w4sq1, # higher = not important
    apprent_unpaid = w1sq11_1,
    apprent_paid = w1sq11_2,
    intern_unpaid = w1sq11_3,
    intern_paid = w1sq11_4,
    psych_12month = w1hq8,
    mental_health_subj = w3hq57, # higher = bad
    inheritance = w4eq10,
    highested = eisced,
    bornincntry = w2sq18a,
    obstacle_firstjob = w3sq71
  )

# filter valid outcome
df = df[df$timetojob %in% 1:4, ]

# clean age
df = df[df$age < 100, ]

# birth year transformation
df$yrbrn0 = df$yrbrn - min(df$yrbrn, na.rm = TRUE)

# recode psych_12month
df$psych_12month[df$psych_12month %in% c(1,2)] = 1
df$psych_12month[df$psych_12month == 3] = 0
df$psych_12month[df$psych_12month == 9] = NA

# other missing recodes
df[df$agegroup35 == 9, "agegroup35"] = NA
df$mental_health_subj[df$mental_health_subj == 9] = NA
df$workimportance_subj[df$workimportance_subj == 9] = NA
df$eduyrs[df$eduyrs >= 77] = NA
df$highested[df$highested >= 55] = NA

# recode bornincntry
df$bornincntry[df$bornincntry == 2] = 0
df$bornincntry[df$bornincntry == 9] = NA

# recode inheritance
df$inheritance[df$inheritance == 2] = 0
df$inheritance[df$inheritance == 9] = NA

# variables to convert to factor
vars = c(
  "during_studies_abroad_studies",
  "during_studies_abroad_work",
  "during_studies_abroad_langcourse",
  "during_studies_abroad_summerschool",
  "agegroup35",
  "apprent_unpaid",
  "apprent_paid",
  "intern_unpaid",
  "intern_paid",
  "inheritance",
  "psych_12month",
  "mental_health_subj",
  "cntry",
  "bornincntry",
  "highested",
  "obstacle_firstjob"
)

df = df %>%
  mutate(across(all_of(vars), ~ factor(.)))

# overview
summary(df[, c("during_studies_abroad_studies", "during_studies_abroad_work",
               "during_studies_abroad_langcourse", "during_studies_abroad_summerschool",
               "agegroup35", "apprent_unpaid", "apprent_paid",
               "intern_unpaid", "intern_paid", "inheritance", "psych_12month",
               "mental_health_subj")])

```

Odds ratio per variable for an ordinal logistic regression model:

```{r modelfitting, echo=FALSE, message=FALSE}
# contrasts(df$cntry) = contr.treatment(nlevels(df$cntry)) # --> austria is reference category
# df$cntry = relevel(df$cntry, ref = "FR")  # France is reference category
contrasts(df$cntry) = contr.sum # --> deviation (sum-to-zero) coding (diff to overall mean)
df$timetojob_fac = factor(df$timetojob,
                      levels = c(1, 2, 3, 4),
                      labels = c("<1 month", "1–6 months", "6–12 months", ">1 year"),
                      ordered = TRUE)
model_cat = polr(timetojob_fac ~ cntry + yrbrn0 +
                  during_studies_abroad_studies +
                  during_studies_abroad_work +
                  during_studies_abroad_langcourse +
                  during_studies_abroad_summerschool +
                  workimportance_subj +
                  apprent_unpaid * inheritance+
                  apprent_paid * inheritance+
                  intern_unpaid * inheritance+
                  intern_paid * inheritance +
                  eduyrs + bornincntry +
                  mental_health_subj, data = df, Hess = TRUE)

#summary(model_cat)
coefs = summary(model_cat)$coefficients

p_values = 2 * pnorm(abs(coefs[, "t value"]), lower.tail = FALSE)

coefs_withp = cbind(exp(coefs[,1]), p_value = p_values, "<0.01" = p_values < 0.01, "<0.05" = p_values < 0.05)

additional_country = exp(-1* sum(coefficients(model_cat)[1:10]))
coefs_withp = rbind(coefs_withp[1:10, ], cntry11 = c(additional_country, NA,NA,NA), coefs_withp[11:nrow(coefs_withp), ])
rownames(coefs_withp)[1:11]  = paste0("cntry", levels(df$cntry))
coefs_withp = coefs_withp[1:32,]

# pretty table output
table_out <- as.data.frame(coefs_withp) %>%
  mutate(
    Variable = rownames(coefs_withp),
    OR = round(V1, 3),
    p_value = signif(p_value, 3),
    Significance = case_when(
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      TRUE ~ ""
    )
  ) %>%
  dplyr::select(Variable, OR, p_value, Significance)
colnames(table_out)[3] = "P-Value"

table_out$Variable <- recode(table_out$Variable,
  "yrbrn0" = "Birth year",
  "eduyrs" = "Years of education",
  "bornincntry1" = "Born in country",
  "workimportance_subj" = "Importance of work",
  
  # Country
  "cntryAT" = "Austria",
  "cntryBE" = "Belgium",
  "cntryCZ" = "Czechia",
  "cntryFI" = "Finland",
  "cntryFR" = "France",
  "cntryGB" = "United Kingdom",
  "cntryHU" = "Hungary",
  "cntryIS" = "Iceland",
  "cntryPL" = "Poland",
  "cntryPT" = "Portugal",
  "cntrySI" = "Slovenia",

  # Mental health
  "mental_health_subj2" = "Mental health: good",
  "mental_health_subj3" = "Mental health: fair",
  "mental_health_subj4" = "Mental health: bad",
  "mental_health_subj5" = "Mental health: very bad",

  # During studies
  "during_studies_abroad_studies1" = "Studied abroad",
  "during_studies_abroad_work1" = "Work/internship abroad",
  "during_studies_abroad_langcourse1" = "Language course abroad",
  "during_studies_abroad_summerschool1" = "Summer school/workshop abroad",

  # Work experience
  "apprent_unpaid1" = "Unpaid apprenticeship",
  "apprent_paid1" = "Paid apprenticeship",
  "intern_unpaid1" = "Unpaid internship",
  "intern_paid1" = "Paid internship",

  # Socioeconomic background
  "inheritance1" = "Expects inheritance",

  # Interactions
  "apprent_unpaid1:inheritance1" = "Unpaid apprenticeship × inheritance",
  "inheritance1:apprent_paid1" = "Paid apprenticeship × inheritance",
  "inheritance1:intern_unpaid1" = "Unpaid internship × inheritance",
  "inheritance1:intern_paid1" = "Paid internship × inheritance"
)
rownames(table_out) = NULL

# variable groups and order
table_out <- table_out %>%
  mutate(
    group = case_when(
      Variable %in% c("Austria","Belgium","Czechia","Finland","France",
                      "United Kingdom","Hungary","Iceland","Poland","Portugal","Slovenia") ~ "Countries",
      
      Variable %in% c("Birth year","Born in country") ~ "Demographics",
      
      grepl("abroad|education|Importance of work", Variable) ~ "Education and studies",
      
      grepl("apprent|intern", Variable) & !grepl("×", Variable) ~ "Work experience",
      
      grepl("Mental health", Variable) ~ "Mental health",
      
      Variable == "Expects inheritance" ~ "Socioeconomic background",
      
      grepl("×", Variable) ~ "Interactions",
      
      TRUE ~ "Other"
    )
  ) %>%
  mutate(
    group = factor(group, levels = c(
      "Countries",
      "Demographics",
      "Education and studies",
      "Work experience",
      "Mental health",
      "Socioeconomic background",
      "Interactions"
    ))
  ) %>%
  arrange(group)

# headers
add_header <- function(group_name) {
  data.frame(
    Variable = group_name,
    OR = NA,
    p_value = NA,
    Significance = "",
    stringsAsFactors = FALSE
  )
}

table_final <- bind_rows(
  add_header("Countries"),
  filter(table_out, group == "Countries"),
  
  add_header("Demographics"),
  filter(table_out, group == "Demographics"),
  
  add_header("Education and studies"),
  filter(table_out, group == "Education and studies"),
  
  add_header("Work experience"),
  filter(table_out, group == "Work experience"),
  
  add_header("Mental health"),
  filter(table_out, group == "Mental health"),
  
  add_header("Socioeconomic background"),
  filter(table_out, group == "Socioeconomic background"),
  
  add_header("Interactions"),
  filter(table_out, group == "Interactions")
)
table_final_clean <- table_final %>%
  mutate(
    `P-Value` = round(`P-Value`, digits = 4)
  ) %>%
  dplyr::select(Variable, OR, `P-Value`, Significance)
options(knitr.kable.NA = '')

kable(table_final_clean,
      col.names = c("Variable", "OR", "P-Value", "Significance"),
      caption = "Determinants of Time to First Job (Odds Ratios)",
      booktabs = TRUE,
      align = "lccc") %>%
  kable_styling(full_width = FALSE) %>%
  add_indent(which(!is.na(table_final_clean$OR)))
  # kable_styling(full_width = FALSE, position = "center") %>%
  # row_spec(which(table_out$`P-Value` < 0.05), bold = TRUE)

```

*Note: Odds ratios above 1 indicate a higher likelihood of longer
job-search duration, while values below 1 indicate faster entry into
employment. Significance levels: \*\* p* $<$ 0.01, \* p $<$ 0.05.

# Interpretation of Results

The outcome variable measures the time until obtaining the first job,
where higher categories indicate a longer transition into employment.

\begin{itemize}
\item An odds ratio greater than 1 indicates a higher likelihood of experiencing longer job-search durations.
\item An odds ratio less than 1 indicates a higher likelihood of faster entry into employment.
\item Statistical significance is evaluated at the 5% and 1% levels.
\end{itemize}

## Country

The results show **substantial cross-country variation** in the
transition from education to employment.\newline

Several countries exhibit significantly **faster transitions**:

\begin{itemize}
\item Austria (OR $=$ 0.77, p $<$ 0.01)
\item Finland (OR $=$ 0.75, p $<$ 0.01)
\item Iceland (OR $=$ 0.32, p $<$ 0.01)
\end{itemize}

In particular, Iceland stands out with a very strong effect, indicating
that individuals there are much less likely to experience prolonged job
search periods.\newline \newline In contrast, some countries show
significantly **slower transitions**:

\begin{itemize}
\item France (OR $=$ 1.82, p $<$ 0.01)
\item Portugal (OR $=$ 1.90, p $<$ 0.01)
\item Poland (OR $=$ 1.43, p $<$ 0.01)
\end{itemize}

This suggests that structural differences between national labour
markets play an important role in shaping early career trajectories.

Other countries (e.g., Belgium, Czech Republic, United Kingdom, Hungary)
do not show statistically significant differences. \newline The
differences between the different countries also become very apparent
when looking at the following plot.

```{r plot1, echo=FALSE, message=FALSE}
cols <- brewer.pal(5, "Blues")[2:5]
tab = prop.table(table(df$timetojob_fac, df$cntry), margin = 2)

df_plot = as.data.frame(tab)
names(df_plot) = c("response", "country", "prop")

df_plot$country <- factor(df_plot$country,
  levels = c("AT","BE","CZ","FI","FR","GB","HU","IS","PL","PT","SI"),
  labels = c("Austria","Belgium","Czechia","Finland","France",
             "United Kingdom","Hungary","Iceland","Poland","Portugal","Slovenia")
)
ggplot(df_plot, aes(x = country, y = prop, fill = factor(response))) +
  geom_bar(stat = "identity") +
  labs(x = "Country",
       y = "Proportion",
       fill = "Time to First Job") +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(family = "serif")
  ) +
  scale_fill_manual(values = cols)
```

## Education and study-related experiences

The results indicate that different types of experiences during studies
have **varying effects**:

\begin{itemize}
\item Working abroad during studies is associated with faster job entry (OR $=$ 0.79, p $<$ 0.05).
\item Attending a language course abroad is associated with slower transitions (OR $=$ 1.42, p $<$ 0.01).
\item Studying abroad and summer schools show no significant effects.
\end{itemize}

These findings suggest that practical work experience **abroad may be
more valuable** for labour market integration than purely academic or
short-term activities.

```{r plot2, echo=FALSE, message=FALSE}
edu_vars <- c(
  "during_studies_abroad_studies",
  "during_studies_abroad_work",
  "during_studies_abroad_langcourse",
  "during_studies_abroad_summerschool",
  "timetojob_fac"
)

df_edu_plot <- df %>%
  dplyr::select(all_of(edu_vars)) %>%
  mutate(
    any_abroad = ifelse(
      during_studies_abroad_studies == 0 &
      during_studies_abroad_work == 0 &
      during_studies_abroad_langcourse == 0 &
      during_studies_abroad_summerschool == 0,
      1, 0
    )
  ) %>%
  mutate(any_abroad = factor(any_abroad)) %>% 
  pivot_longer(cols = c("during_studies_abroad_studies",
                        "during_studies_abroad_work",
                        "during_studies_abroad_langcourse",
                        "during_studies_abroad_summerschool",
                        "any_abroad"),
               names_to = "activity",
               values_to = "did_activity")

df_edu_plot = df_edu_plot[df_edu_plot$did_activity == 1,]

tab = prop.table(table(df_edu_plot$timetojob_fac, df_edu_plot$activity), margin = 2)
df_plot = as.data.frame(tab)
names(df_plot) = c("response", "experience", "prop")

df_plot$experience <- factor(df_plot$experience,
  levels = c("during_studies_abroad_studies",
             "during_studies_abroad_work",
             "during_studies_abroad_langcourse",
             "during_studies_abroad_summerschool",
             "any_abroad"),
  labels = c("Studied abroad",
             "Work/internship abroad",
             "Language course abroad",
             "Summer school/workshop abroad",
             "None")
)
ggplot(df_plot, aes(x = experience, y = prop, fill = factor(response))) +
  geom_bar(stat = "identity") +
  labs(x = "Abroad Experience",
       y = "Proportion",
       fill = "Time to First Job") +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(family = "serif")
  ) +
  scale_fill_manual(values = cols)
```

## Apprenticeships and internships

Work experience plays an important but nuanced role:

\begin{itemize}
\item Paid apprenticeships (OR $=$ 0.77, p $<$ 0.01) and unpaid apprenticeships (OR $=$ 0.77, p $<$ 0.01) are both associated with faster transitions into employment.
\item Paid internships show a weak, non-significant tendency toward faster entry (OR $=$ 0.86).
\item Unpaid internships show no significant effect.
\end{itemize}

This indicates that **apprenticeships are particularly effective** in
facilitating labour market entry, while internships appear less
consistently beneficial.

```{r plot3, echo=FALSE, message=FALSE}
edu_vars <- c(
  "apprent_paid",
  "apprent_unpaid",
  "intern_paid",
  "intern_unpaid",
  "timetojob_fac"
)

df_edu_plot <- df %>%
  dplyr::select(all_of(edu_vars)) %>%
  mutate(
    any_abroad = ifelse(
      apprent_paid == 0 &
      apprent_unpaid == 0 &
      intern_paid == 0 &
      intern_unpaid == 0,
      1, 0
    )
  ) %>%
  mutate(any_abroad = factor(any_abroad)) %>% 
  pivot_longer(cols = c("apprent_paid",
                        "apprent_unpaid",
                        "intern_paid",
                        "intern_unpaid",
                        "any_abroad"),
               names_to = "activity",
               values_to = "did_activity")

df_edu_plot = df_edu_plot[df_edu_plot$did_activity == 1,]


tab = prop.table(table(df_edu_plot$timetojob_fac, df_edu_plot$activity), margin = 2)
df_plot = as.data.frame(tab)
names(df_plot) = c("response", "experience", "prop")

df_plot$experience <- factor(df_plot$experience,
  levels = c("apprent_paid",
             "apprent_unpaid",
             "intern_paid",
             "intern_unpaid",
             "any_abroad"),
  labels = c("Paid Apprenticeship",
             "Unpaid Apprenticeship",
             "Paid Internship",
             "Unpaid Internship",
             "Neither")
)
ggplot(df_plot, aes(x = experience, y = prop, fill = factor(response))) +
  geom_bar(stat = "identity") +
  labs(x = "Apprenticeships and internships",
       y = "Proportion",
       fill = "Time to First Job") +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(family = "serif")
  ) +
  scale_fill_manual(values = cols)
```

## Socioeconomic background

The expectation of receiving an **inheritance does not show a
significant** direct effect (OR $=$ 1.02). The interaction terms between
inheritance and work experience are also not statistically significant.
This suggests that the effects of apprenticeships and internships on job
entry do not differ substantially depending on expected inheritance.
\newline

However, another important factor is:

\begin{itemize}
\item Being born in the country is associated with substantially faster job entry (OR $=$ 0.56, p $<$ 0.01).
\end{itemize}

This suggests that structural advantages, such as familiarity with the
local labour market or fewer barriers, play a key role. This implies
that, people **not born in the country**, such as immigrants, have
**more difficulty entering the labour market quickly**.

```{r plot4, echo=FALSE, message=FALSE}
edu_vars <- c(
  "bornincntry",
  "timetojob_fac"
)

df_edu_plot <- df %>%
  dplyr::select(all_of(edu_vars)) 


tab = prop.table(table(df_edu_plot$timetojob_fac, df_edu_plot$bornincntry), margin = 2)
df_plot = as.data.frame(tab)
names(df_plot) = c("response", "bornincntry", "prop")

df_plot$bornincntry <- factor(df_plot$bornincntry,
  levels = c(1, 0),
  labels = c("Yes", "No")
)
ggplot(df_plot, aes(x = bornincntry, y = prop, fill = factor(response))) +
  geom_bar(stat = "identity") +
  labs(x = "Born in Country",
       y = "Proportion",
       fill = "Time to First Job") +
  scale_y_continuous(labels = percent) +
  theme_minimal() + theme(text = element_text(family = "serif")) +
  scale_fill_manual(values = cols)
```

## Mental health

Mental health shows one of the **strongest and most consistent
patterns** in the model. \newline

Compared to individuals with very good mental health (which serves as
reference category):

\begin{itemize}
\item Moderate difficulties (OR $\approx$ 1.39–1.64) are associated with longer job search durations
\item Severe difficulties (OR $=$ 3.08, p $<$ 0.01) are associated with substantially longer transitions
\end{itemize}

This indicates a clear gradient effect, where **worsening mental
health** is strongly linked to **increasing difficulty in entering
employment**.

```{r plot5, echo=FALSE, message=FALSE}
edu_vars <- c(
  "mental_health_subj",
  "timetojob_fac"
)

df_edu_plot <- df %>%
  dplyr::select(all_of(edu_vars)) 

tab = prop.table(table(df_edu_plot$timetojob_fac, df_edu_plot$mental_health_subj), margin = 2)
df_plot = as.data.frame(tab)
names(df_plot) = c("response", "mental", "prop")

df_plot$experience <- factor(df_plot$mental,
  levels = c(1:5),
  labels = c("Very good",
             "Good",
             "Fair",
             "Bad",
             "Very bad")
)
# ggplot(df_plot, aes(x = experience, y = prop, fill = factor(response))) +
#   geom_bar(stat = "identity") +
#   labs(x = "Mental Health",
#        y = "Proportion",
#        fill = "Time to First Job") +
#   scale_y_continuous(labels = percent) +
#   theme_minimal() + theme(text = element_text(family = "serif")) +
#   scale_fill_manual(values = cols)
ggplot(df_plot, aes(x=as.numeric(as.character(mental)), y=prop, fill=response)) +
  geom_area() +
  labs(x = "Mental Health",
       y = "Proportion",
       fill = "Time to First Job") +   
  scale_x_continuous(
    breaks = 1:5,
    labels = c("Very good", "Good", "Fair", "Bad", "Very bad")
  ) +
  theme_minimal() + theme(text = element_text(family = "serif")) +
  scale_fill_manual(values = cols)
```

## Other individual characteristics

\begin{itemize}
\item Age (year of birth): Younger individuals tend to experience slightly longer transitions (OR $=$ 1.02, p $<$ 0.01), indicating that the job search has become more difficult throughout time.
\item Years of education: More education is associated with a slightly longer transition (OR $=$ 1.02, p $<$ 0.01), possibly reflecting longer job search for better matches.
\item Importance of work: No significant effect.
\end{itemize}

```{r plot6, echo=FALSE, message=FALSE}
df$age_group = cut(df$age,
                    breaks = c(15, 20, 25, 30, 35, 40),
                    labels = c("15–20", "21–25", "26–30", "31–35", "36–40"))

tab = prop.table(table(df$timetojob_fac, df$age_group), margin = 2)

df_plot = as.data.frame(tab)
names(df_plot) = c("response", "age_group", "prop")
ggplot(df_plot, aes(x = age_group, y = prop, fill = factor(response))) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Age group", y = "Proportion", fill = "Response") +
  theme_minimal() +
  theme(text = element_text(family = "serif")) +
  scale_fill_manual(values = cols)

df_plot = df %>%
  group_by(age, timetojob_fac) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(age) %>%
  mutate(prop = n / sum(n))

ggplot(df_plot, aes(x = age, y = prop, color = timetojob_fac)) +
  labs(x = "Age", y = "Proportion", color = "Response") +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  theme(text = element_text(family = "serif")) +
  scale_color_manual(values = cols) 
```

# Policy Implications and Recommandations

This analysis shows that the transition from education to employment is
shaped not only by individual choices, but by structural factors such as
labour market institutions, access to work experience, and mental
health.

## Mental health is a key barrier {.unnumbered}

Poor mental health is strongly associated with delayed entry into
employment. \newline $\rightarrow$ Integrate mental health support into
education and employment policies, especially during the transition
phase.

## Labour market structures matter {.unnumbered}

There are large differences between countries in how quickly young
people find jobs. \newline $\rightarrow$ Learn from high-performing
systems (e.g. Iceland) and adapt best practices at the national level.

## Inequalities affect opportunities {.unnumbered}

Being born in the country significantly improves employment chances,
indicating structural barriers for some groups. \newline $\rightarrow$
Provide targeted support for disadvantaged youth, including migrants and
those with fewer resources.

## Apprenticeships are highly effective {.unnumbered}

Structured work experience, especially apprenticeships, significantly
improves early employment outcomes, while internships are less
consistent. \newline $\rightarrow$ Expand and strengthen apprenticeship
systems and improve the quality and regulation of internships.
